### 2016-02-03 results/tables/build_table1.Rmd
Build Table 1

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=1.0in
---

\captionsetup{labelformat=empty}

```{r table1, results='asis', echo=FALSE, warning=FALSE, eval=TRUE}
require(xtable, quietly=TRUE)
options(xtable.sanitize.text.function=identity)
options(xtable.sanitize.colnames.function=identity)

pretty_region <- c("V4"="V4", "V1V3"="V1-V3", "V3V5"="V3-V5", "V1V5"="V1-V5", "V1V6"="V1-V6", "V1V9"="V1-V9")
regions <- names(pretty_region)

mock_error <- read.table(file="data/process/mock.error.report")
mock_error_nochim <- mock_error[mock_error$numparents==1,]
mock_error_good <- mock_error_nochim[mock_error_nochim$error <= 0.1,]
n_seqs <- aggregate(mock_error_good$error, by=list(mock_error_good$region), length)
rownames(n_seqs) <- n_seqs$Group.1
n_seqs <- n_seqs[regions,]

getSloppyRevComp <- function(sequence){
	rev.sequence <- paste(rev(unlist(strsplit(sequence, ""))), collapse="")
	revcomp.sequence <- chartr("ATGCRYSWKMBDHVN", "TACG...........", rev.sequence)
	return(revcomp.sequence)
}

getPos <- function(primer, sequence=ecoli, forward=TRUE){
	if(forward==FALSE){
		primer <- getSloppyRevComp(primer)
	}
	primer <- gsub("[^ATGC]", ".", primer)

	pos <- regexpr(primer, sequence)

	if(forward==FALSE){
		pos <- pos + nchar(primer) -1
	}
	return(pos)

}

#taken from... http://www.ncbi.nlm.nih.gov/nuccore/NC_022648.1?report=fasta&from=4564120&to=4565661
ecoli <- "AAATTGAAGAGTTTGATCATGGCTCAGATTGAACGCTGGCGGCAGGCCTAACACATGCAAGTCGAACGGTAACAGGAAGCAGCTTGCTGCTTTGCTGACGAGTGGCGGACGGGTGAGTAATGTCTGGGAAACTGCCTGATGGAGGGGGATAACTACTGGAAACGGTAGCTAATACCGCATAACGTCGCAAGACCAAAGAGGGGGACCTTCGGGCCTCTTGCCATCGGATGTGCCCAGATGGGATTAGCTAGTAGGTGGGGTAACGGCTCACCTAGGCGACGATCCCTAGCTGGTCTGAGAGGATGACCAGCCACACTGGAACTGAGACACGGTCCAGACTCCTACGGGAGGCAGCAGTGGGGAATATTGCACAATGGGCGCAAGCCTGATGCAGCCATGCCGCGTGTATGAAGAAGGCCTTCGGGTTGTAAAGTACTTTCAGCGGGGAGGAAGGGAGTAAAGTTAATACCTTTGCTCATTGACGTTACCCGCAGAAGAAGCACCGGCTAACTCCGTGCCAGCAGCCGCGGTAATACGGAGGGTGCAAGCGTTAATCGGAATTACTGGGCGTAAAGCGCACGCAGGCGGTTTGTTAAGTCAGATGTGAAATCCCCGGGCTCAACCTGGGAACTGCATCTGATACTGGCAAGCTTGAGTCTCGTAGAGGGGGGTAGAATTCCAGGTGTAGCGGTGAAATGCGTAGAGATCTGGAGGAATACCGGTGGCGAAGGCGGCCCCCTGGACGAAGACTGACGCTCAGGTGCGAAAGCGTGGGGAGCAAACAGGATTAGATACCCTGGTAGTCCACGCCGTAAACGATGTCGACTTGGAGGTTGTGCCCTTGAGGCGTGGCTTCCGGAGCTAACGCGTTAAGTCGACCGCCTGGGGAGTACGGCCGCAAGGTTAAAACTCAAATGAATTGACGGGGGCCCGCACAAGCGGTGGAGCATGTGGTTTAATTCGATGCAACGCGAAGAACCTTACCTGGTCTTGACATCCACGGAAGTTTTCAGAGATGAGAATGTGCCTTCGGGAACCGTGAGACAGGTGCTGCATGGCTGTCGTCAGCTCGTGTTGTGAAATGTTGGGTTAAGTCCCGCAACGAGCGCAACCCTTATCCTTTGTTGCCAGCGGTCCGGCCGGGAACTCAAAGGAGACTGCCAGTGATAAACTGGAGGAAGGTGGGGATGACGTCAAGTCATCATGGCCCTTACGACCAGGGCTACACACGTGCTACAATGGCGCATACAAAGAGAAGCGACCTCGCGAGAGCAAGCGGACCTCATAAAGTGCGTCGTAGTCCGGATTGGAGTCTGCAACTCGACTCCATGAAGTCGGAATCGCTAGTAATCGTGGATCAGAATGCCACGGTGAATACGTTCCCGGGCCTTGTACACACCGCCCGTCACACCATGGGAGTGGGTTGCAAAAGAAGTAGGTAGCTTAACCTTCGGGAGGGCGCTTACCACTTTGTGATTCATGACTGGGGTGAAGTCGTAACAAGGTAACCGTAGGGGAACCTGCGGTTGGATCACCTCCTTA"

oligos <- read.table(file="data/references/pacbio.oligos", stringsAsFactors=F)[1:6,]
rownames(oligos) <- oligos$V4
oligos <- oligos[,c("V2", "V3")]
colnames(oligos) <- c("forward", "reverse")
oligos <- oligos[regions,]

start <- sapply(oligos[,1], getPos)
end <- sapply(oligos[,2], getPos, forward=FALSE)

coordinates <- paste(start, end, sep="-")
length <- end - start - nchar(oligos[,1]) - nchar(oligos[,2]) + 1
composite <- cbind(region = pretty_region, oligos, coorinates=coordinates, length=format(length, digits=0), nseqs=format(n_seqs$x, big.mark=','))

header <- c("Region", "\\bigcell{c}{ Forward }", "\\bigcell{c}{ Reverse }", "\\bigcell{c}{\\textit{E. coli} \\\\ Coordinates}", "\\bigcell{c}{Amplicon \\\\ Length}", "\\bigcell{c}{Sequences \\\\ (N)}")

header <- paste0("\\textbf{", header, "}")

addtorow <- list(pos=list(0),
                command=paste0(paste(header, collapse=" & "), "\\\\"))

composite <- xtable(composite)

caption <- "\\textbf{Table 1. Summary of the primer pairs used to generate the 16S rRNA gene fragment fragments and the characteristics of each region.}"

composite <- xtable(composite, caption = caption, label="")

align(composite) <- "ccllccc"

print(composite,
    add.to.row=addtorow,
	include.rownames=FALSE,
	include.colnames=FALSE,
	hline.after = c(-1, 0, nrow(composite)),
    comment=FALSE,
    caption.placement = "top",
    table.placement="!ht",
	size="small"
)
```