### 2016-02-12 results/tables/build_table2.Rmd
Highlander comments

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=1.0in
---

\captionsetup{labelformat=empty}

```{r table1, results='asis', echo=FALSE, warning=FALSE, eval=TRUE}
require(xtable, quietly=TRUE)
options(xtable.sanitize.text.function=identity)
options(xtable.sanitize.colnames.function=identity)

pretty_region <- c("V4"="V4", "V1V3"="V1-V3", "V3V5"="V3-V5", "V1V5"="V1-V5", "V1V6"="V1-V6", "V1V9"="V1-V9")
region_order <- names(pretty_region)

mock_error <- read.table(file="data/process/error_summary.tsv", header=T, stringsAsFactors=F)
regions <- mock_error[mock_error$method=="basic","region"]
basic <- mock_error[mock_error$method=="basic","error"]
predicted <- mock_error[mock_error$method=="pred","error"]
precluster <- mock_error[mock_error$method=="precluster","error"]
percent_remaining <- 100-mock_error[mock_error$method=="precluster","nseqs_red"]
error_data <- data.frame(cbind(basic, predicted, precluster, percent_remaining))
rownames(error_data) <- regions
error_data <- error_data[region_order,]
error_data <- format(error_data, nsmall=2L, digits=2)

sobs_data <- read.table(file="data/process/sobs_table.tsv", header=T)
sobs_data <- sobs_data[region_order,-c(1,7)]
sobs_data <- format(sobs_data, nsmall=1L, digits=2)

composite <- cbind(error_data, sobs_data)
rownames(composite) <- pretty_region

#Error rate (%) for:
#% reads remaining from basic
#Average no. of OTUsa

header <- c("Basic", "Predicted", "Precluster", "", "Mock\\textsuperscript{b}", "Mock\\textsuperscript{c}", "Soil", "Mouse", "Human")

header <- paste0("\\textbf{", header, "}")
composite <- rbind(" " = header, composite)

addtorow <- list(pos=list(0,1),
                command=c('\\multirow{3}{*}{\\textbf{Region}} & \\multicolumn{3}{c}{\\textbf{Error rate following (\\%)}} & \\multirow{1}{*}{\\textbf{\\bigcell{c}{Reads \\\\ remaining (\\%)}}} & \\multicolumn{5}{c}{\\textbf{Average no. of OTUs\\textsuperscript{a}}} \\\\ \\cmidrule(r){2-4}	\\cmidrule(r){6-10}', '\\hline '))

caption <- "\\textbf{Table 2. Summary of the error rates and number of observed OTUs for each region.}"

composite <- xtable(composite, caption = caption, label="")

align(composite) <- "cccccccccc"
print(composite,
    add.to.row=addtorow,
	include.rownames=TRUE,
	include.colnames=FALSE,
	hline.after = c(-1, nrow(composite)),
    comment=FALSE,
    caption.placement = "top",
    table.placement="!ht",
	size="small",
	NA.string="ND"
)
```

^a^ The number of OTUs is based on rarefaction of each sample to 1,000 sequences per sample; cells labeled ND reflect samples that did not have at least 1,000 sequences.

^b^ Number of OTUs in the mock community when all chimeras were removed; in the absence of chimeras and sequencing errors, there should be 19 OTUs for all three regions.

^c^ Number of OTUs in the mock community when chimeras were removed using UCHIME.