### 2016-02-18 submission/Schloss_PacBio16S_PeerJ_2016.tex
Create TeX intermediate files

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{\textbf{Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA
sequencing system}}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle

\vspace{35mm}

\textbf{Running title:} 16S rRNA genes sequencing with PacBio

\vspace{15mm}

\textbf{Authors:} Patrick D. Schloss\textsuperscript{1\#}, Matthew L.
Jenior\textsuperscript{1}, Charles C. Koumpouras\textsuperscript{1},
Sarah L. Westcott\textsuperscript{1}, and Sarah K.
Highlander\textsuperscript{2}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

1. Department of Microbiology and Immunology, 1500 W. Medical Center,
University of Michigan, Ann Arbor, MI 48109

2. Genomic Medicine, J. Craig Venter Institute, 4120 Capricorn Lane, La
Jolla, CA 92307

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

Over the past 10 years, microbial ecologists have largely abandoned
sequencing 16S rRNA genes by the Sanger sequencing method and have
instead adopted highly parallelized sequencing platforms. These new
platforms, such as 454 and Illumina's MiSeq, have allowed researchers to
obtain millions of high quality, but short sequences. The result of the
added sequencing depth has been significant improvements in experimental
design. The tradeoff has been the decline in the number of full-length
reference sequences that are deposited into databases. To overcome this
problem, we tested the ability of the PacBio Single Molecule, Real-Time
(SMRT) DNA sequencing platform to generate sequence reads from the 16S
rRNA gene. We generated sequencing data from the V4, V3-V5, V1-V3,
V1-V5, V1-V6, and V1-V9 variable regions from within the 16S rRNA gene
using DNA from a synthetic mock community and natural samples collected
from human feces, mouse feces, and soil. The mock community allowed us
to assess the actual sequencing error rate and how that error rate
changed when different curation methods were applied. We developed a
simple method based on sequence characteristics and quality scores to
reduce the observed error rate for the V1-V9 region from 0.69 to
0.027\%. This error rate is comparable to what has been observed for the
shorter reads generated by 454 and Illumina's MiSeq sequencing
platforms. Although the per base sequencing cost is still significantly
more than that of MiSeq, the prospect of supplementing reference
databases with full-length sequences from organisms below the limit of
detection from the Sanger approach is exciting.

\textbf{\emph{Keywords:}} Microbial ecology, bioinformatics, sequencing
error

\newpage

\subsection{Introduction}\label{introduction}

Advances in sequencing technologies over the past 10 years have
introduced considerable advances to the field of microbial ecology.
Clone-based Sanger sequencing of the 16S rRNA gene has largely been
replaced by various platforms produced by 454/Roche (e.g. Sogin et al.
(2006)), Illumina (e.g. Gloor et al. (2010)), and IonTorrent (e.g.
Jünemann et al. (2012)). It was once common to sequence fewer than 100
16S rRNA gene sequences from several samples using the Sanger approach
(e.g. Mccaig et al. (1999)). Now it is common to generate thousands of
sequences from each of several hundred samples (Consortium, 2012). The
advance in throughput has come at the cost of read length. Sanger
sequencing regularly generated 800 nt per read and because the DNA was
cloned, it was possible to obtain multiple reads per fragment to yield a
full-length sequence from a representative single molecule. At
approximately \$8 (US) per sequencing read, most researchers have
effectively decided that full-length sequences are not worth the
increased cost relative to the cost of more recently developed
approaches. There is still a clear need to generate high-throughput
full-length sequence reads that are of sufficient quality that they can
be used as references for analyses based on obtaining short sequence
reads.

Historically, all sequencing platforms were created to primarily perform
genome sequencing. When sequencing a genome, it is assumed that the same
base of DNA will be sequenced multiple times and the consensus of
multiple sequence reads is used to generate contigs. Thus, an individual
base call may have a high error rate, but the consensus sequence will
have a low error rate. To sequence the 16S rRNA gene researchers use
conserved primers to amplify a sub-region from within the gene that is
isolated from many organisms. Because the fragments are not cloned, it
is not possible to obtain high sequence coverage from the same DNA
molecule using these platforms. To reduce sequencing error rates it has
become imperative to develop stringent sequence curation and denoising
algorithms (Schloss, Gevers \& Westcott, 2011; Kozich et al., 2013).
There has been a tradeoff between read length, number of reads per
sample, and the error rate. For instance, we recently demonstrated that
using the Illumina MiSeq and the 454 Titanium platforms the raw error
rate varies between 1 and 2\% (Schloss, Gevers \& Westcott, 2011; Kozich
et al., 2013). Yet, it was possible to obtain error rates below 0.02\%
by adopting various denoising algorithms; however, the resulting
fragments were only 250-nt long. In the case of 454 Titanium, extending
the length of the fragment introduces length-based errors and in the
case of the Illumina MiSeq, increasing the length of the fragment
reduces the overlap between the read pairs reducing the ability of each
read to mutually reduce the sequencing error. Inadequate denoising of
sequencing reads can have many negative effects including limited
ability to identify chimeras (Haas et al., 2011; Edgar et al., 2011) and
inflation of alpha- and beta-diversity metrics (Kunin et al., 2010; Huse
et al., 2010; Schloss, Gevers \& Westcott, 2011; Kozich et al., 2013).
Illumina's MiSeq plaform enjoys widespread use in the field because of
the ability to sequence 15-20 million fragments that can be distributed
across hundreds of samples for less than \$5000 (US).

As these sequencing platforms have grown in popularity, there has been a
decline in the number of full-length 16S rRNA genes being deposited into
GenBank that could serve as references for sequence classification,
phylogenetic analyses, and primer and probe design. This is particularly
frustrating since the technologies have significantly improved our
ability to detect and identify novel populations for which we lack
full-length reference sequences. A related problem is the perceived
limitation that the short reads generated by the 454 and Illumina
platforms cannot be reliably classified to the genus or species level.
Previous investigators have utilized simulations to demonstrate that
increased read lengths usually increase the accuracy and sensitivity of
classification against reference databases (Wang et al., 2007; Liu et
al., 2008; Werner et al., 2011). There is clearly a need to develop
sequencing technologies that will allow researchers to generate high
quality full-length 16S rRNA gene sequences in a high throughput manner.

New advances in single molecule sequencing technologies are being
developed to address this problem. One approach uses a random barcoding
strategy to fragment, sequence, and assemble full-length amplicons using
Illumina's HiSeq platform (Miller et al., 2013; Burke \& Darling, 2014).
Although the algorithms appear to have minimized the risk of assembly
chimeras, it is unclear what the sequencing error rate is by this
approach. An alternative is the use of single molecule technologies that
offer read lengths that are thousands of bases long. Although the Oxford
Nanopore Technology has been used to sequence 16S rRNA genes
(Benítez-Páez, Portune \& Sanz, 2016), the the platform produced by
Pacific Biosciences (PacBio) has received wider attention for this
application (Fichot \& Norman, 2013; Mosher et al., 2013, 2014; Schloss
et al., 2015; Singer et al., 2016). The PacBio Single Molecule,
Real-Time (SMRT) DNA Sequencing System ligates hairpin adapters
(i.e.~SMRTbells) to the ends of double-stranded DNA. Although the DNA
molecule is linear, the adapters effectively circularize the DNA
allowing the sequencing polymerase to process around the molecule
multiple times (Au et al., 2012). According to Pacific Biosciences the
platform is able to generate median read lengths longer than 8 kb with
the P6-C4 chemistry; however, the single pass error rate is
approximately 15\%. Given the circular nature of the DNA fragment, the
full read length can be used to cover the DNA fragment multiple times
resulting in a reduced error rate. Therefore, one should be able to
obtain multiple coverage of the full 16S rRNA gene at a reduced error
rate.

Despite the opportunity to potentially generate high-quality full-length
sequences, it is surprising that the Pacific Biosciences platform has
not been more widely adopted for sequencing 16S rRNA genes. Previous
studies utilizing the technology have removed reads with mismatched
primers and barcodes, ambiguous base calls, and low quality scores
(Fichot \& Norman, 2013) or screened sequences based on the predicted
error rate (Singer et al., 2016). Others have utilized the platform
without describing the bioinformatic pipeline that was utilized (Mosher
et al., 2013, 2014). The only study to report the error rate of the
platform for sequencing 16S rRNA genes with a mock community used the
P4-C2 chemistry and obtained an error rate of 0.32\%, which is 16-fold
higher than has been observed using the MiSeq or 454 platforms (Schloss,
Gevers \& Westcott, 2011; Kozich et al., 2013). In the current study, we
assessed the quality of data generated by the PacBio sequencer using the
improved P6-C4 chemistry and on-sequencer data processing. The goal was
to determine whether this strategy could fill the need for generating
high-quality, full-length sequence data on par with other platforms. We
hypothesized that by modulating the 16S rRNA gene fragment length we
could alter the read depth and obtain reads longer than are currently
available by the 454 and Illumina platforms but with the same quality.
To test this hypothesis, we developed a sequence curation pipeline that
was optimized by reducing the sequencing error rate of a mock bacterial
community with known composition. The resulting pipeline was then
applied to 16S rRNA gene fragments that were isolated from soil and
human and mouse feces.

\subsection{Materials and Methods}\label{materials-and-methods}

\textbf{\emph{Community DNA.}} We utilized genomic DNA isolated from
four communities. These same DNA extracts were previously used to
develop an Illumina MiSeq-based sequencing strategy (Kozich et al.,
2013). Briefly, we used a ``Mock Community'' composed of genomic DNA
from 21 bacterial strains: \emph{Acinetobacter baumannii} ATCC 17978,
\emph{Actinomyces odontolyticus} ATCC 17982, \emph{Bacillus cereus} ATCC
10987, \emph{Bacteroides vulgatus} ATCC 8482, \emph{Clostridium
beijerinckii} ATCC 51743, \emph{Deinococcus radiodurans} ATCC 13939,
\emph{Enterococcus faecalis} ATCC 47077, \emph{Escherichia coli} ATCC
70096, \emph{Helicobacter pylori} ATCC 700392, \emph{Lactobacillus
gasseri} ATCC 33323, \emph{Listeria monocytogenes} ATCC BAA-679,
\emph{Neisseria meningitidis} ATCC BAA-335, \emph{Porphyromonas
gingivalis} ATCC 33277, \emph{Propionibacterium acnes} DSM 16379,
\emph{Pseudomonas aeruginosa} ATCC 47085, \emph{Rhodobacter sphaeroides}
ATCC 17023, \emph{Staphylococcus aureus} ATCC BAA-1718,
\emph{Staphylococcus epidermidis} ATCC 12228, \emph{Streptococcus
agalactiae} ATCC BAA-611, \emph{Streptococcus mutans} ATCC 700610,
\emph{Streptococcus pneumoniae} ATCC BAA-334. The mock community DNA is
available through BEI resources (v3.1, HM-278D). Genomic DNAs from the
three other communities were obtained using the MO BIO PowerSoil DNA
extraction kit. The human and mouse fecal samples were obtained using
protocols that were reviewed and approved by the University Committee on
Use and Care of Animals (Protocol \#PRO00004877) and the Institutional
Review Board at the University of Michigan (Protocol \#HUM00057066). The
human stool donor provided informed consent.

\textbf{\emph{Library generation and sequencing.}} The DNAs were each
amplified in triplicate using barcoded primers targeting the V4, V1-V3,
V3-V5, V1-V5, V1-V6, and V1-V9 variable regions (Table 1). The primers
were synthesized so that the 5' end of the forward and reverse primers
were each tagged with paired 16-nt symmetric barcodes
(\url{https://github.com/PacificBiosciences/Bioinformatics-Training/wiki/Barcoding-with-SMRT-Analysis-2.3})
to allow multiplexing of samples within a single sequencing run. Methods
describing PCR, amplicon cleanup, and pooling were described previously
(Kozich et al., 2013). The SMRTbell adapters were ligated onto the PCR
products and the libraries were sequenced by Pacific Biosciences using
the P6-C4 chemistry on a PacBio RS II SMRT DNA Sequencing System.
Diffusion Loading was used for regions V4, V1-V3, and V3-V5 and MagBead
loading was used for regions V1-V5, V1-V6, and V1-V9. Each region was
sequenced separately using movies ranging in length between 180 and 360
minutes. The sequences were processed using pbccs (v.3.0.1;
\url{https://github.com/PacificBiosciences/pbccs}), which generates
predicted error rates using a proprietary algorithm.

\textbf{\emph{Data analysis.}} All sequencing data were curated using
mothur (v1.36)(Schloss et al., 2009) and analyzed using the R
programming language (R Core Team, 2016). The raw data can be obtained
from the Sequence Read Archive at NCBI under accession SRP051686, which
are associated with BioProject PRJNA271568. This accession and
bioproject also contain data from the same samples sequenced using the
P4-C2 chemistry. Several specific features were incorporated into mothur
to facilitate the analysis of PacBio sequence data. First, because
non-ambiguous base calls are assigned to Phred quality scores of zero,
the consensus fastq files were parsed so that scores of zero were
interpreted as corresponding to an ambiguous base call (i.e.~N) in the
fastq.info command using the pacbio=T option. Second, because the
consensus sequence can be generated in the forward and reverse
complement orientations, a checkorient option was added to the trim.seqs
command in order to identify the proper orientation. These features were
incorporated into mothur v.1.30. Because chimeric molecules can be
generated during PCR and would artificially inflate the sequencing
error, it was necessary to remove these data prior to assessing the
error rate. Because we knew the true sequences for the strains in the
mock community we could calculate all possible chimeras between strains
in the mock community (\emph{in silico} chimeras). If a sequence read
was 3 or more nucleotides more similar to an \emph{in silico} chimera
than it was to a non-chimeric reference sequence, it was classified as a
chimera and removed from further consideration. Identification of
\emph{in silico} chimeras and calculation of sequencing error rates was
performed using the seq.error command in mothur (Schloss, Gevers \&
Westcott, 2011). \emph{De novo} chimera detection was also performed on
the mock and other sequence data using the abundance-based algorithm
implemented in UCHIME (Edgar et al., 2011). Sequences sequences were
aligned against a SILVA-based reference alignment (Pruesse et al., 2007)
using a profile-based aligner (Schloss, 2009) and were classified
against the SILVA (v123) (Pruesse et al., 2007), RDP (v10)(Cole et al.,
2013), and greengenes (v13\_8\_99)(Werner et al., 2011) reference
taxonomies using a negative Bayesian classifier implemented within
mothur (Wang et al., 2007). Sequences were assigned to operational
taxonomic units using the average neighbor clustering algorithm with a
3\% distance threshold (Schloss \& Westcott, 2011). Detailed methods
including this paper as an R markdown file are available as a public
online repository
(\url{http://github.com/SchlossLab/Schloss_PacBio16S_PeerJ_2016}).

\subsection{Results and Discussion}\label{results-and-discussion}

\textbf{\emph{The PacBio error profile and a basic sequence curation
procedure.}} To build a sequence curation pipeline, we first needed to
characterize the error rate associated with sequencing the 16S rRNA
gene. Using the consensus sequence obtained from at least 3 sequencing
passes of each fragment, we observed an average sequencing error rate of
0.65\%. Insertions, deletions, and substitutions accounted for 31.2,
17.9, and 50.9\% of the errors, respectively. The substitution errors
were equally likely and all four bases were equally likely to cause
insertion errors. Interestingly, guanines (39.4\%) and adenines (24.3\%)
were more likely to be deleted than cytosines (18.3\%) or thymidines
(18.0\%). The PacBio quality values varied between 2 and 93.
Surprisingly, the percentage of base calls that had the maximum quality
value did not vary among correct base calls (80.5\%), substitutions
(80.0\%), and insertions (80.4\%). It was not possible to use the
individual base quality scores could not be used to screen sequence
quality as has been possible in past studies using the Phred-based
quality scores that accompany data generated using the 454 and Illumina
technologies. We did observe a strong correlation between our observed
error rate and the predicted error rate as calculated by the PacBio
software (Pearson's R: -0.67; Figure 1A).

We established a simple curation procedure by culling any sequence that
had a string of the same base repeated more than 8 times or did not
start and end at the expected alignment coordinates for that region of
the 16S rRNA gene. This reduced the experiment-wide error rate from 0.68
to 0.65\%. This basic procedure resulted in the removal of between 0.714
(V1-V3) and 9.47 (V1-V9)\% of the reads (Table 2). Although the
percentage of reads removed increased with the length of the fragment,
there was no obvious relationship between fragment length and error rate
(Figure 2).

\textbf{\emph{Identifying correlates of increased sequencing error.}} In
contrast to the 454 and Illumina-based platforms where the sequencing
quality decays with length, the consensus sequencing approach employed
by the PacBio sequencer is thought to generate a uniform distribution of
errors. This makes it impossible to simply trim sequences to high
quality regions. Therefore, we sought to identify characteristics within
sequences that would allow us to identify and remove those sequences
with errors using three different approaches. First, we hypothesized
that errors in the barcode and primer would be correlated with the error
rate for the entire sequence. We observed a strong relationship between
the number of mismatches to the barcodes and primers and the error rate
of the rest of the sequence fragment (Figure 1B). Although allowing no
mismatches to the barcodes and primers yielded the lowest error rate,
that stringent criterion removed a large fraction of the reads from the
dataset. Allowing at most one mismatch only marginally increased the
error rate while retaining more sequences in the dataset (Figure 2).
Second, we hypothesized that increased sequencing coverage should yield
lower error rates. We found that once we had obtained 10-fold coverage
of the fragments, the error rate did not change appreciably (Figure 1C).
When we compared the error rates of reads with at least 10-fold coverage
to those with less coverage, we reduced the error rate by 8.48 to
37.08\% (Figure 2). Third, based on the observed correlation between the
predicted and observed error rates, we sought to identify a minimum
predicted error rate that would allow us to reduce the observed error
rate. The average observed error rate for sequences with predicted error
rates between 0.01 and 0.10\% was linear. We decided to use a threshold
of 0.01\% because a large number of sequence reads were lost when we
used a smaller threshold. When we used this threshold, we were able to
reduce the error rate by 51.4 to 70.0\% (Figure 2). Finally, we
quantified the effect of combining filters. We found that any
combination of filters that included the predicted error rate threshold
had the most significant impact on reducing the observed error rate.
Furthermore, the inclusion of the mismatch and coverage filters had a
negligible impact on error rates, but had a significant impact on the
number of sequences included in the analysis. For instance, among the
V1-V9 data, requiring sequences to have a predicted error rate less than
0.01\% resulted in a 69.2\% reduction in error and resulted in the
removal of 53.5\% of the sequences. Adding the mismatch or coverage
filter had no effect on the reduction of error, but resulted in the
removal of 56.2 and 56.6 \% of the sequences, respectively. Use of all
filters had no impact on the reduction in the observed error rate, but
resulted in the removal of 59.1\% of the sequences. The remainder of
this paper only uses sequences with a predicted error rate less than
0.01\%.

\textbf{\emph{Pre-clustering sequences to further reduce sequencing
noise.}} Previously, we implemented a pre-clustering algorithm where
sequences were sorted by their abundance in decreasing order and rare
sequences are clustered with a more abundant sequence if the rare
sequences have fewer mismatches than a defined threshold when compared
to the more abundant sequence (Huse et al., 2010; Schloss, Gevers \&
Westcott, 2011). The recommended threshold was a 1-nt difference per
100-nt of sequence data. For example, the threshold for 250 bp fragment
from the V4 region would be 2 nt or 14 for the 1458 bp V1-V9 fragments.
This approach removes residual PCR and sequencing errors while not
overwhelming the resolution needed to identify OTUs that are based on a
3\% distance threshold. The tradeoff of this approach is that one would
be unable to differentiate V1-V9 sequences that truly differed by less
than 14 nt. When we applied this approach to our PacBio data, we
observed a reduction in the error rate between 33.0 (V1-V3) and 48.7\%
(V1-V9). The final error rates varied between 0.02 (V1-V5) and 0.2\%
(V4). The full-length (i.e.~V1-V9) fragments had an error rate of 0.03\%
(Figure 2; Table 2), this is similar to what we have previously observed
using the 454 and Illumina MiSeq platforms (0.02\%)(Schloss, Gevers \&
Westcott, 2011; Kozich et al., 2013).

\textbf{\emph{Effects of error rates on OTU assignments.}} The
sequencing error rate is known to affect the number of OTUs that are
observed (Schloss, Gevers \& Westcott, 2011). For each region, we
determined that if there were no chimeras or PCR or sequencing errors,
then we would expect to find 19 OTUs. When we achieved perfect chimera
removal, but allowed for PCR and sequencing errors, we observed between
0.5 (V1-V9) and 14.4 (V4) extra OTUs (Table 2). The range in the number
of extra OTUs was largely explained by the sequencing error rate
(Pearson's R=1.0). Next, we determined the number of OTUs that were
observed when we used UCHIME to identify chimeric sequences. Under these
more realistic conditions, we observed between 8.2 (V1-V5) and 29.9 (V4)
extra OTUs. Finally, we calculated the number of OTUs in the soil,
mouse, and human samples using the same pipeline with chimera detection
and removal based on the UCHIME algorithm. Surprisingly, there was not a
clear relationship across sample type and region. Again, we found that
there was a strong correlation between the number of observed OTUs and
the error rate for the mouse (R=0.95) and human samples (R=0.60). These
results underscore the effect of sequencing error on the inflation of
the number of observed OTUs.

\textbf{\emph{Classification varies by region, environment, and
database.}} We classified all of the sequence data we generated using
the naïve Bayesian classifier using the RDP, SILVA, and greengenes
reference taxonomies (Figure 3). In general, increasing the length of
the region improved the ability to assign the sequence to a genus or
species. Interestingly, each of the samples we analyzed varied in the
ability to assign its sequences to the depth of genus or species.
Furthermore, the reference database that did the best job of classifying
the sequences varied by sample type. For example, the SILVA reference
did the best for the human feces and soil samples and the RDP did the
best for the mouse feces samples. An advantage of the greengenes
database is that it contains information for 2,514 species-level
lineages for 11\% of the reference sequences; the other databases only
provided taxonomic data to the genus level. There was a modest
association between the length of the fragment and the ability to
classify sequences to the species-level for the human samples; there was
no such association for the mouse and soil samples. In fact, at most
6.2\% of the soil sequences and 4.3\% of the mouse sequences could be
classified to a species. These results indicate that the ability to
classify sequences to the genus or species level is a function of read
length, sample type, and the reference database.

\textbf{\emph{Sequencing errors are not random.}} Above, we described
that although there was no obvious bias in the substitution or insertion
rate, we did observe that guanines and adenines were more likely to be
deleted than cytosines or thymidines. This lack of randomness in the
error profile suggested that there might be a systematic non-random
distribution of the errors across the sequences. This would manifest
itself by the creation of duplicate sequences with the same error. We
identified all of the mock community sequences that had a 1-nt
difference to the true sequence (Figure 4). For these three regions,
between 70.7 and 88.9 of the sequences with 1-nt errors were only
observed once. We found that the frequency of the most abundant 1-nt
error paralleled the number of sequences. Surprisingly, the same 1-nt
error appeared 1,954 times (0.02\%) in the V1-V6 mock data and another
1-nt error appeared 1,070 times (0.03\%) in the V1-V9 mock data.
Contrary to previous reports (Carneiro et al., 2012; Koren et al.,
2012), these results indicate that reproducible errors occur with the
PacBio sequencing platform and that they can be quite abundant. Through
the use of the pre-clustering step described above these 1-nt errors
would be ameliorated; however, this result indicates that caution should
be used when attempting to use fine-scale OTU definitions.

\subsection{Conclusions}\label{conclusions}

The various sequencing platforms that are available to microbial
ecologists are able to fill unique needs and have their own strengths
and weaknesses. For sequencing the 16S rRNA gene, the 454 platform is
able to generate a moderate number of high-quality 500-nt sequence
fragments (error rates below 0.02\%) (Schloss, Gevers \& Westcott, 2011)
and the MiSeq platform is able to generate a large number of
high-quality 250-nt sequence fragments (error rates below 0.02\%)
(Kozich et al., 2013). The promise of the PacBio sequencing platform was
the generation of high-quality near full-length sequence fragments. As
we have shown in this study, it is possible to generate near full-length
sequences with error rates that are slightly higher, but comparable to
the other platforms (i.e.~0.03\%). With the exception of the V4 region
(0.2\%), the error rates were less than 0.07\%. When we considered the
shorter V4 region, which is similar in length to what is sequenced by
the MiSeq platform, the error rates we observed with the PacBio platform
were nearly 8-fold higher than what has previously been reported on the
other platforms. It was unclear why these shorter reads had such a high
error rate relative to the other regions. At this point, the primary
limitation of generating full-length sequences on the PacBio platform is
the cost of generating the data and accessibility to the sequencers.

The widespread adoption of the 454 and MiSeq platforms and decrease in
the use of Sanger sequencing for the 16S rRNA gene has resulted in a
decrease in the generation of the full-length reference sequences that
are needed for performing phylogenetic analyses and designing lineage
specific PCR primers and fluorescent \emph{in situ} hybridization (FISH)
probes. It remains to be determined whether the error rates we observed
for full-length sequences are prohibitive for these applications. We can
estimate the distribution of errors assuming that the errors follow a
binomial distribution along the length of the 1,500 nt gene with the
error rate that we achieved from the V1-V9 mock community data prior to
pre-clustering the sequences, which was 0.2\%. Under these conditions
one would expect 4.3\% of the sequences to have no errors and 50\% of
the sequences would have at least 3 errors. After applying the
pre-clustering denoising step, the error rate drops by 7.7-fold to
0.03\%. With this error rate, we would expect 66.3\% of the sequences to
have no sequencing errors. The cost of the reduced error rate is the
loss of resolution among closely related sequences.

Full-length sequences are frequently seen as a panacea to overcome the
limitations of taxonomic classifications. The ability to classify each
of our sample types benefited from the generation of full-length
sequences. It was interesting that the benefit varied by sample type and
database. For example, using the mouse libraries, the ability to
classify each of the regions differed by less than 5\% when classifying
against the SILVA and greengenes databases. The effect of the database
that was used was also interesting. The RDP database outperformed the
other databases for the mouse samples and the SILVA database
outperformed the others for the human and soil samples. The three
databases were equally effective for classifying the mock community.
Finally, since only the greengenes database provided species-level
information for its reference sequences it was the only database that
allowed for resolution of species-level classification. The sequences
from the mouse and soil libraries were not effectively classified to the
species level (all less than 10\%). In contrast, classification of the
human libraries resulted in more than 40\% of the sequences being
classified to a genus, regardless of the region. These data demonstrate
that for the samples we analyzed, the length of the sequence fragment
was not as significant a factor in classification as the choice of
database.

The development of newer sequencing technologies continue to advance and
there is justifiable excitement to apply these technologies to sequence
the 16S rRNA gene. Although it is clearly possible to generate
sequencing data from these various platforms, it is critical that we
assess the platforms for their ability to generate high quality data and
the particular niche that the new approach will fill. With this in mind,
it is essential that researchers utilize mock communities as part of
their experimental design so that they can quantify their error rates.
The ability to generate near full-length 16S rRNA gene sequences is an
exciting advance that will hopefully expand our ability to improve the
characterization of microbial communities.

\subsection{Acknowledgements}\label{acknowledgements}

The Genomic DNA from Microbial Mock Community A (Even, Low
Concentration, v3.1, HM-278D) was obtained through the NIH Biodefense
and Emerging Infections Research Resources Repository, NIAID, NIH as
part of the Human Microbiome Project.

\newpage

\subsection{Figures}\label{figures}

\textbf{Figure 1. Summary of errors in data generated using PacBio
sequencing platform to sequence various regions within the 16S rRNA
gene.} The predicted error rate using PacBio's sequence analysis
algorithm correlated well with the observed error rate (Pearson's R:
-0.67; A). Because of the large number of sequences, we randomly
selected 5\% of the data to show in panel A. The sequencing error rate
of the amplified gene fragments increased with mismatches to the
barcodes and primers (B). The sequencing error rate declined with
increased sequencing coverage; however, increasing the sequencing depth
beyond 10-fold coverage had no meaningful effect on the sequencing error
rate (C). The scale of they y-axis in B and C are the same.

\textbf{Figure 2. Change in error rate (A) and the percentage of
sequences that were retained (B) when using various sequence curation
methods.} The condition that was used for downstream analyses is
indicated by the star. The plotted numbers represent the region that was
sequenced. For example ``19'' represents the data for the V1-V9 region.

\textbf{Figure 3. Percentage of unique sequences that could be
classified.} Classifications were performed using taxonomy references
curated from the RDP, SILVA, or greengenes databases for the four types
of samples that were sequenced across the six regions from the 16S rRNA
gene. Only the greengenes taxonomy reference provided species-level
information.

\textbf{Figure 4. Percentage of 1-nt variants that occurred up to ten
times.} Sequences that were 1 nt different from the mock community
reference sequences were counted to determine the number of times each
variant appeared by region within the 16S rRNA gene.

\newpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-Au2012}{}
Au KF., Underwood JG., Lee L., Wong WH. 2012. Improving PacBio long read
accuracy by short read alignment. \emph{PLoS ONE} 7:e46679. DOI:
\href{https://doi.org/10.1371/journal.pone.0046679}{10.1371/journal.pone.0046679}.

\hypertarget{ref-BentezPez2016}{}
Benítez-Páez A., Portune KJ., Sanz Y. 2016. Species-level resolution of
16S rRNA gene amplicons sequenced through the MinION portable nanopore
sequencer. \emph{GigaScience} 5. DOI:
\href{https://doi.org/10.1186/s13742-016-0111-z}{10.1186/s13742-016-0111-z}.

\hypertarget{ref-Burke2014}{}
Burke C., Darling AE. 2014. Resolving microbial microdiversity with high
accuracy full length 16S rRNA illumina sequencing. DOI:
\href{https://doi.org/10.1101/010967}{10.1101/010967}.

\hypertarget{ref-Carneiro2012}{}
Carneiro MO., Russ C., Ross MG., Gabriel SB., Nusbaum C., DePristo MA.
2012. Pacific biosciences sequencing technology for genotyping and
variation discovery in human data. \emph{BMC Genomics} 13:375. DOI:
\href{https://doi.org/10.1186/1471-2164-13-375}{10.1186/1471-2164-13-375}.

\hypertarget{ref-Cole2013}{}
Cole JR., Wang Q., Fish JA., Chai B., McGarrell DM., Sun Y., Brown CT.,
Porras-Alfaro A., Kuske CR., Tiedje JM. 2013. Ribosomal database
project: Data and tools for high throughput rRNA analysis. \emph{Nucleic
Acids Research} 42:D633--D642. DOI:
\href{https://doi.org/10.1093/nar/gkt1244}{10.1093/nar/gkt1244}.

\hypertarget{ref-HMP2012}{}
Consortium THM. 2012. Structure, function and diversity of the healthy
human microbiome. \emph{Nature} 486:207--214. DOI:
\href{https://doi.org/10.1038/nature11234}{10.1038/nature11234}.

\hypertarget{ref-Edgar2011}{}
Edgar RC., Haas BJ., Clemente JC., Quince C., Knight R. 2011. UCHIME
improves sensitivity and speed of chimera detection.
\emph{Bioinformatics} 27:2194--2200. DOI:
\href{https://doi.org/10.1093/bioinformatics/btr381}{10.1093/bioinformatics/btr381}.

\hypertarget{ref-Fichot2013}{}
Fichot EB., Norman RS. 2013. Microbial phylogenetic profiling with the
pacific biosciences sequencing platform. \emph{Microbiome} 1:10. DOI:
\href{https://doi.org/10.1186/2049-2618-1-10}{10.1186/2049-2618-1-10}.

\hypertarget{ref-Gloor2010}{}
Gloor GB., Hummelen R., Macklaim JM., Dickson RJ., Fernandes AD.,
MacPhee R., Reid G. 2010. Microbiome profiling by illumina sequencing of
combinatorial sequence-tagged PCR products. \emph{PLoS ONE} 5:e15406.
DOI:
\href{https://doi.org/10.1371/journal.pone.0015406}{10.1371/journal.pone.0015406}.

\hypertarget{ref-Haas2011}{}
Haas BJ., Gevers D., Earl AM., Feldgarden M., Ward DV., Giannoukos G.,
Ciulla D., Tabbaa D., Highlander SK., Sodergren E., Methe B., DeSantis
TZ., Petrosino JF., Knight R., Birren BW. 2011. Chimeric 16S rRNA
sequence formation and detection in sanger and 454-pyrosequenced PCR
amplicons. \emph{Genome Research} 21:494--504. DOI:
\href{https://doi.org/10.1101/gr.112730.110}{10.1101/gr.112730.110}.

\hypertarget{ref-Huse2010}{}
Huse SM., Welch DM., Morrison HG., Sogin ML. 2010. Ironing out the
wrinkles in the rare biosphere through improved OTU clustering.
\emph{Environmental Microbiology} 12:1889--1898. DOI:
\href{https://doi.org/10.1111/j.1462-2920.2010.02193.x}{10.1111/j.1462-2920.2010.02193.x}.

\hypertarget{ref-Junemann2012}{}
Jünemann S., Prior K., Szczepanowski R., Harks I., Ehmke B., Goesmann
A., Stoye J., Harmsen D. 2012. Bacterial community shift in treated
periodontitis patients revealed by ion torrent 16S rRNA gene amplicon
sequencing. \emph{PLoS ONE} 7:e41606. DOI:
\href{https://doi.org/10.1371/journal.pone.0041606}{10.1371/journal.pone.0041606}.

\hypertarget{ref-Koren2012}{}
Koren S., Schatz MC., Walenz BP., Martin J., Howard JT., Ganapathy G.,
Wang Z., Rasko DA., McCombie WR., Jarvis ED., Phillippy AM. 2012. Hybrid
error correction and de novo assembly of single-molecule sequencing
reads. \emph{Nat Biotechnol} 30:693--700. DOI:
\href{https://doi.org/10.1038/nbt.2280}{10.1038/nbt.2280}.

\hypertarget{ref-Kozich2013}{}
Kozich JJ., Westcott SL., Baxter NT., Highlander SK., Schloss PD. 2013.
Development of a dual-index sequencing strategy and curation pipeline
for analyzing amplicon sequence data on the MiSeq illumina sequencing
platform. \emph{Applied and Environmental Microbiology} 79:5112--5120.
DOI: \href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\hypertarget{ref-Kunin2010}{}
Kunin V., Engelbrektson A., Ochman H., Hugenholtz P. 2010. Wrinkles in
the rare biosphere: Pyrosequencing errors can lead to artificial
inflation of diversity estimates. \emph{Environmental Microbiology}
12:118--123. DOI:
\href{https://doi.org/10.1111/j.1462-2920.2009.02051.x}{10.1111/j.1462-2920.2009.02051.x}.

\hypertarget{ref-Liu2008}{}
Liu Z., DeSantis TZ., Andersen GL., Knight R. 2008. Accurate taxonomy
assignments from 16S rRNA sequences produced by highly parallel
pyrosequencers. \emph{Nucleic Acids Research} 36:e120--e120. DOI:
\href{https://doi.org/10.1093/nar/gkn491}{10.1093/nar/gkn491}.

\hypertarget{ref-Mccaig1999}{}
Mccaig AE., Glover LA., James., Prosser I. 1999. Molecular analysis of
bacterial community structure and diversity in unimproved and improved
upland grass pastures. \emph{Appl Environ Microbiol} 65:1721--1730.

\hypertarget{ref-Miller2013}{}
Miller CS., Handley KM., Wrighton KC., Frischkorn KR., Thomas BC.,
Banfield JF. 2013. Short-read assembly of full-length 16S amplicons
reveals bacterial diversity in subsurface sediments. \emph{PLoS ONE}
8:e56018. DOI:
\href{https://doi.org/10.1371/journal.pone.0056018}{10.1371/journal.pone.0056018}.

\hypertarget{ref-Mosher2013}{}
Mosher JJ., Bernberg EL., Shevchenko O., Kan J., Kaplan LA. 2013.
Efficacy of a 3rd generation high-throughput sequencing platform for
analyses of 16S rRNA genes from environmental samples. \emph{Journal of
Microbiological Methods} 95:175--181. DOI:
\href{https://doi.org/10.1016/j.mimet.2013.08.009}{10.1016/j.mimet.2013.08.009}.

\hypertarget{ref-Mosher2014}{}
Mosher JJ., Bowman B., Bernberg EL., Shevchenko O., Kan J., Korlach J.,
Kaplan LA. 2014. Improved performance of the PacBio SMRT technology for
16S rDNA sequencing. \emph{Journal of Microbiological Methods}
104:59--60. DOI:
\href{https://doi.org/10.1016/j.mimet.2014.06.012}{10.1016/j.mimet.2014.06.012}.

\hypertarget{ref-Pruesse2007}{}
Pruesse E., Quast C., Knittel K., Fuchs BM., Ludwig W., Peplies J.,
Glockner FO. 2007. SILVA: A comprehensive online resource for quality
checked and aligned ribosomal RNA sequence data compatible with ARB.
\emph{Nucleic Acids Research} 35:7188--7196. DOI:
\href{https://doi.org/10.1093/nar/gkm864}{10.1093/nar/gkm864}.

\hypertarget{ref-R2016}{}
R Core Team. 2016. R: A language and environment for statistical
computing.

\hypertarget{ref-Schloss2009a}{}
Schloss PD., Westcott SL., Ryabin T., Hall JR., Hartmann M., Hollister
EB., Lesniewski RA., Oakley BB., Parks DH., Robinson CJ., Sahl JW.,
Stres B., Thallinger GG., Horn DJV., Weber CF. 2009. Introducing mothur:
Open-source, platform-independent, community-supported software for
describing and comparing microbial communities. \emph{Applied and
Environmental Microbiology} 75:7537--7541. DOI:
\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

\hypertarget{ref-Schloss2009b}{}
Schloss PD. 2009. A high-throughput DNA sequence aligner for microbial
ecology studies. \emph{PLoS ONE} 4:e8230. DOI:
\href{https://doi.org/10.1371/journal.pone.0008230}{10.1371/journal.pone.0008230}.

\hypertarget{ref-Schloss2015}{}
Schloss PD., Westcott SL., Jenior ML., Highlander; SK. 2015. Sequencing
16S rRNA gene fragments using the pacBio sMRT dNA sequencing system.
DOI:
\href{https://doi.org/10.7287/peerj.preprints.778v1}{10.7287/peerj.preprints.778v1}.

\hypertarget{ref-Schloss2011b}{}
Schloss PD., Westcott SL. 2011. Assessing and improving methods used in
operational taxonomic unit-based approaches for 16S rRNA gene sequence
analysis. \emph{Applied and Environmental Microbiology} 77:3219--3226.
DOI: \href{https://doi.org/10.1128/aem.02810-10}{10.1128/aem.02810-10}.

\hypertarget{ref-Schloss2011a}{}
Schloss PD., Gevers D., Westcott SL. 2011. Reducing the effects of PCR
amplification and sequencing artifacts on 16S rRNA-based studies.
\emph{PLoS ONE} 6:e27310. DOI:
\href{https://doi.org/10.1371/journal.pone.0027310}{10.1371/journal.pone.0027310}.

\hypertarget{ref-Singer2016}{}
Singer E., Bushnell B., Coleman-Derr D., Bowman B., Bowers RM., Levy A.,
Gies EA., Cheng J-F., Copeland A., Klenk H-P., Hallam SJ., Hugenholtz
P., Tringe SG., Woyke T. 2016. High-resolution phylogenetic microbial
community profiling. \emph{The ISME Journal}. DOI:
\href{https://doi.org/10.1038/ismej.2015.249}{10.1038/ismej.2015.249}.

\hypertarget{ref-Sogin2006}{}
Sogin ML., Morrison HG., Huber JA., Welch DM., Huse SM., Neal PR.,
Arrieta JM., Herndl GJ. 2006. Microbial diversity in the deep sea and
the underexplored ``rare biosphere''. \emph{Proceedings of the National
Academy of Sciences} 103:12115--12120. DOI:
\href{https://doi.org/10.1073/pnas.0605127103}{10.1073/pnas.0605127103}.

\hypertarget{ref-Wang2007}{}
Wang Q., Garrity GM., Tiedje JM., Cole JR. 2007. Naive bayesian
classifier for rapid assignment of rRNA sequences into the new bacterial
taxonomy. \emph{Applied and Environmental Microbiology} 73:5261--5267.
DOI: \href{https://doi.org/10.1128/aem.00062-07}{10.1128/aem.00062-07}.

\hypertarget{ref-Werner2011}{}
Werner JJ., Koren O., Hugenholtz P., DeSantis TZ., Walters WA., Caporaso
JG., Angenent LT., Knight R., Ley RE. 2011. Impact of training sets on
classification of high-throughput bacterial 16s rRNA gene surveys.
\emph{The ISME Journal} 6:94--103. DOI:
\href{https://doi.org/10.1038/ismej.2011.82}{10.1038/ismej.2011.82}.

\end{document}