### 2016-03-16 submission/table_2.tex
Correct rules generating TeX files for Tables 1 and 2

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage{multirow}

\pagenumbering{gobble}

\usepackage{array}

\newcommand{\bigcell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}

\usepackage{rotating}

\usepackage{caption}

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle

\captionsetup{labelformat=empty}

\begin{table}[!ht]
\centering
\caption{\textbf{Table 2. Summary of the error rates and number of observed OTUs for each region.}} 
\label{}
{\small
\begin{tabular}{cccccccccc}
  \hline
  \multirow{3}{*}{\textbf{Region}} & \multicolumn{3}{c}{\textbf{Error rate following (\%)}} & \multirow{1}{*}{\textbf{\bigcell{c}{Reads \\ remaining (\%)}}} & \multicolumn{5}{c}{\textbf{Average no. of OTUs\textsuperscript{a}}} \\ \cmidrule(r){2-4} \cmidrule(r){6-10}  & \textbf{Basic} & \textbf{Predicted} & \textbf{Precluster} & \textbf{} & \textbf{Mock\textsuperscript{b}} & \textbf{Mock\textsuperscript{c}} & \textbf{Soil} & \textbf{Mouse} & \textbf{Human} \\ 
   \hline V4 & 0.77 & 0.35 & 0.158 & 62.58 & 33.4 & 48.9 & 629.4 & 212.6 & 116.9 \\ 
  V1-V3 & 0.56 & 0.27 & 0.068 & 67.01 & 20.5 & 27.7 & 726.0 & 107.3 & 106.4 \\ 
  V3-V5 & 0.74 & 0.22 & 0.035 & 63.15 & 21.1 & 31.2 &    NA & 112.3 &  83.8 \\ 
  V1-V5 & 0.57 & 0.21 & 0.019 & 59.29 & 19.8 & 27.2 & 694.4 &  65.9 &  84.9 \\ 
  V1-V6 & 0.60 & 0.21 & 0.031 & 59.90 & 20.0 & 46.5 & 693.0 &  75.8 &  89.4 \\ 
  V1-V9 & 0.68 & 0.21 & 0.027 & 51.33 & 19.5 & 40.4 &    NA & 102.5 & 117.3 \\ 
   \hline
\end{tabular}
}
\end{table}

\textsuperscript{a} The number of OTUs is based on rarefaction of each
sample to 1,000 sequences per sample; cells labeled ND reflect samples
that did not have at least 1,000 sequences.

\textsuperscript{b} Number of OTUs in the mock community when all
chimeras were removed; in the absence of chimeras and sequencing errors,
there should be 19 OTUs for all three regions.

\textsuperscript{c} Number of OTUs in the mock community when chimeras
were removed using UCHIME.

\end{document}